---
title: "Clark-Izzi-ADA-Homework-4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork) # for plotting
library(scales) # for log y axis scale
library(collapse) # for `fmode()` to find mode
library(BBmisc) # for %nin%
library(devtools)
remotes::install_github("joachim-gassen/tidycovid19")
library(tidycovid19)
```

```{r}
merged <- download_merged_data(cached = TRUE)
merged <- merged %>%
  group_by(country, iso3c) %>%
  arrange(country, iso3c, date) %>%
  mutate(
    daily_confirmed = confirmed - lag(confirmed, n = 1),
    daily_deaths = deaths - lag(deaths, n = 1),
    daily_recovered = recovered - lag(recovered, n = 1)
  ) %>%
  mutate(
    daily_confirmed = replace_na(daily_confirmed, 0),
    daily_deaths = replace_na(daily_deaths, 0),
    daily_recovered = replace_na(daily_recovered, 0)
  ) %>%
  ungroup() %>%
  arrange(country, iso3c, date)

add_world1 <- merged %>%
  group_by(date) %>%
  arrange(date) %>%
  summarize(
    country = "World", iso3c = NA, confirmed = sum(confirmed),
    deaths = sum(deaths), recovered = sum(recovered),
    timestamp = fmode(timestamp)
  ) %>%
  mutate(
    daily_confirmed = confirmed - lag(confirmed, n = 1),
    daily_deaths = deaths - lag(deaths, n = 1),
    daily_recovered = recovered - lag(recovered, n = 1)
  ) %>%
  mutate(
    daily_confirmed = replace_na(daily_confirmed, 0),
    daily_deaths = replace_na(daily_deaths, 0),
    daily_recovered = replace_na(daily_recovered, 0)
  ) %>%
  ungroup() %>%
  arrange(country, iso3c, date)

add_world2 <- merged %>%
  group_by(country, iso3c) %>%
  summarize(
    population = fmode(population),
    land_area_skm = fmode(land_area_skm),
    timestamp = fmode(timestamp)
  ) %>%
  ungroup() %>%
  summarize(
    country = "World", iso3c = NA,
    population = sum(population, na.rm = TRUE),
    land_area_skm = sum(land_area_skm, na.rm = TRUE)
  ) %>%
  mutate(pop_density = population / land_area_skm)

add_world <- left_join(add_world1, add_world2, by = c("country", "iso3c"))
merged <- bind_rows(merged, add_world)

cv_data <- pivot_longer(merged,
  cols = c(
    "confirmed", "deaths", "recovered",
    "daily_confirmed", "daily_deaths", "daily_recovered"
  ),
  names_to = "variable", values_to = "cases"
) %>%
  arrange(country, variable, date) %>%
  rename(area = land_area_skm, density = pop_density) %>%
  mutate(rate = cases / population * 10^6)

cv_summary <- function(d, country_list = "World",
                       plot = TRUE, facet = "country",
                       status = c("confirmed", "deaths", "recovered")) {

  # based on `wes_palettes()` GrandBudapest1, IsleofDogs1, IsleofDogs2
  # from the {wesanderson} package
  my_palette <- c(
    "#5B1A18", "#FD6467", "#F1BB7B", "#D67236",
    "#0F0D0E", "#9986A5", "#79402E", "#CCBA72", "#D9D0D3", "#8D8680",
    "#EAD3BF", "#AA9486", "#B6854D", "#39312F", "#1C1718"
  )

  if (facet == "country") {
    fill <- "variable"
    n <- length(unique(d$variable)) / 2
    # need only half of unique # of variable (3)
  }

  if (facet == "variable") {
    fill <- "country"
    n <- length(country_list)
    # need number of countries
  }

  if ("All" %in% country_list) {
    country_list <- unique(d$country)
    country_list <- setdiff(country_list, "World")
  }

  if ("World" %in% country_list) {
    d <- d %>% filter(country %in% country_list)

    totals <- d %>%
      group_by(variable) %>%
      summarize(
        country = "World",
        cases = max(cases),
        population = max(population),
        area = max(area),
        density = max(density),
        rate = max(rate, na.rm = TRUE),
        on = max(date)
      ) %>%
      select(country, variable, cases, population, area, density, rate, on) %>%
      arrange(variable) %>%
      ungroup()
  }

  if ("World" %nin% country_list) {
    d <- d %>% filter(country %in% country_list)
    totals <- d %>%
      group_by(country, variable) %>%
      summarize(
        cases = max(cases),
        population = max(population),
        area = max(area),
        density = max(density),
        rate = max(rate, na.rm = TRUE),
        on = max(date),
        gdp_capita = fmode(gdp_capita),
        income = fmode(income),
        life_expectancy = fmode(life_expectancy),
        max_sd = max(soc_dist),
        max_mr = max(mov_rest)
      ) %>%
      select(
        country, variable, cases, population, area, density, rate,
        gdp_capita, income, life_expectancy, max_sd, max_mr, on
      ) %>%
      arrange(country, variable) %>%
      ungroup()
  }

  if (plot == TRUE) {
    cc <- filter(d, variable %in% status)
    cum_cases_plot <- ggplot(
      data = cc,
      # use the tidy evaluation pronoun .data to slice the chosen fill
      # variable from the data frame
      aes(
        x = date, y = cases + 1, color = .data[[fill]],
        fill = .data[[fill]]
      )
    ) +
      geom_point(size = 0.5) +
      geom_line() +
      # use the tidy evaluation pronoun .data to slice the chosen facet_wrap
      # variable from the data frame
      facet_wrap(~ .data[[facet]], ncol = 5) +
      xlab("Date") +
      ylab("Log Cumulative Cases") +
      scale_y_log10(
        breaks = trans_breaks("log10", function(x) 10^x),
        labels = trans_format("log10", math_format(10^.x))
      ) +
      scale_color_manual(
        aesthetics = c("color", "fill"),
        name = NULL, values = my_palette
      )

    dc <- filter(d, variable %in% paste0("daily_", status))
    daily_cases_plot <- ggplot(
      data = dc,
      aes(
        x = date, y = cases, color = .data[[fill]],
        fill = .data[[fill]]
      )
    ) +
      geom_point(size = 0.5) +
      geom_line() +
      facet_wrap(~ .data[[facet]], ncol = 5) +
      xlab("Date") +
      ylab("Daily Cases") +
      scale_color_manual(
        aesthetics = c("color", "fill"),
        name = NULL, values = my_palette
      )
  }

  if (plot == TRUE) {
    return(list(
      totals = totals,
      cum_cases_plot = cum_cases_plot,
      daily_cases_plot = daily_cases_plot
    ))
  } else {
    return(list(totals = totals))
  }
}
```
cv_summary takes vector country_list=, Boolean plot=, character arg facet = (faceted by country or variable name), vector of covid status variables status =. Outputs 3 elements: plot of cumulative cases, plot of daily cases, and table that contains summary of variables for each country.

## CHALLENGE 1
Plot global data on confirmed coronavirus infections, deaths, and recoveries. (Faceted by variable rather than country, although it doesn't really matter in this case.)
```{r}
cv_summary(cv_data, country_list = "World", plot = TRUE, facet = "variable")
```

## CHALLENGE 2
Plot data on confirmed coronavirus infections, deaths, and recoveries for G7 countries plus China, Russia, and Iran. Facet your plots first by “country” and then by “variable”.
*** Is there a way to facet by both or is country first fine?
```{r}
cv_summary(cv_data, country_list = c("US", "United Kingdom", "Canada", "France", "Germany", "Italy", "Japan", "China", "Russia", "Iran"), plot = TRUE, facet = "country")
```

## CHALLENGE 3
Summary data for ALL countries in the dataset, and then filter this returned dataset to only those countries with populations of over 1 million, storing this dataset as a tibble d.
How many countries does this tibble include? 918
```{r}
summ <- cv_summary(cv_data, country_list = "All", plot = FALSE)
d <- summ$totals # since tibble is outputted in list of one element
d <- filter(d, population > 1000000)
nrow(d)
```

## CHALLENGE 4
Filter d to generate two additional tibbles, overall and daily.
Which 10 countries have experienced the highest overall rate of confirmed cases? Which 10 countries have experienced the highest single-day rate of confirmed cases? See below:
```{r}
# OVERALL CONFIRMED CASES
overall <- filter(d, variable == "confirmed") %>% arrange(desc(rate))
## Top 10:
top_n(overall, 10, rate)$country
# DAILY CONFIRMED
daily <- filter(d, variable == "daily_confirmed") %>% arrange(desc(rate))
## Top 10:
top_n(daily, 10, rate)$country
```

## CHALLENGE 5 - FIGURE OUT WHY lm on logs won't run
Run a linear model to evaluate how the overall infection rate (rate) is related to the variables population density (density), population size (population), gross domestic product per capita (gdp_capita), and overall income level (income). In doing so, you should run exploratory visualizations to see whether or not the four numeric variables should be transformed.

Based on the full model, what predictors variables have slopes significantly different from zero?
```{r}
# DENSITY
lm(rate ~ density, data = d)
# untransformed
ggplot(d, aes(x = density, y = rate)) + geom_point() + geom_smooth(method = "lm")
# log-log transformed
lm(log(rate) ~ log(density), data = d) 
ggplot(d, aes(x = log(density), y = log(rate))) + geom_point() + geom_smooth(method = "lm")
lm(rate ~ population, data = d)
lm(rate ~ gdp_capita, data = d)
lm(rate ~ income, data = d)
```

## CHALLENGE 6
Run stepwise selection using AIC to evaluate whether the full model or a nested, simpler model is preferred. (Module 22 for AIC, 24 for pseudoR2)
What is the best model (based on AIC) of the possible ones involving these 4 predictors?
What are the “pseudo-R2" values associated with the full and “best” models?


Repeat this modeling process to evaluate what combination of explanatory variables best maximum daily infection rate. Are the important predictors the same? What additional or different variables are included?

## CHALLENGE 7
To the best model you determined in CHALLENGE 6 for predicting the maximum daily infection rate, add in the maximum social distancing (max_sd) and maximum movement restriction (max_mr) score per country. Do either of these additional variables improve the model significantly?

## CHALLENGE 8

