---
title: "Clark-Izzi-ADA-Homework-4"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork) # for plotting
library(scales) # for log y axis scale
library(collapse) # for `fmode()` to find mode
library(BBmisc) # for %nin%
library(devtools)
library(tidycovid19)
remotes::install_github("joachim-gassen/tidycovid19")
merged <- download_merged_data(cached = TRUE)
merged <- merged %>%
  group_by(country, iso3c) %>%
  arrange(country, iso3c, date) %>%
  ## new code to replace NAs with zeros
  mutate(
    confirmed = ifelse(is.na(confirmed), 0, confirmed),
    deaths = ifelse(is.na(deaths), 0, deaths),
    recovered = ifelse(is.na(recovered), 0, recovered)
  ) %>%
  ## end of new code
  mutate(
    daily_confirmed = confirmed - lag(confirmed, n = 1),
    daily_deaths = deaths - lag(deaths, n = 1),
    daily_recovered = recovered - lag(recovered, n = 1)
  ) %>%
  mutate(
    daily_confirmed = replace_na(daily_confirmed, 0),
    daily_deaths = replace_na(daily_deaths, 0),
    daily_recovered = replace_na(daily_recovered, 0)
  ) %>%
  ungroup() %>%
  arrange(country, iso3c, date)

add_world1 <- merged %>%
  group_by(date) %>%
  arrange(date) %>%
  summarize(
    country = "World", iso3c = NA,
    confirmed = sum(confirmed, na.rm = TRUE),
    deaths = sum(deaths, na.rm = TRUE),
    recovered = sum(recovered, na.rm = TRUE),
    timestamp = fmode(timestamp)
  ) %>%
  mutate(
    daily_confirmed = confirmed - lag(confirmed, n = 1),
    daily_deaths = deaths - lag(deaths, n = 1),
    daily_recovered = recovered - lag(recovered, n = 1)
  ) %>%
  mutate(
    daily_confirmed = replace_na(daily_confirmed, 0),
    daily_deaths = replace_na(daily_deaths, 0),
    daily_recovered = replace_na(daily_recovered, 0)
  ) %>%
  ungroup() %>%
  arrange(country, iso3c, date)

add_world2 <- merged %>%
  group_by(country, iso3c) %>%
  summarize(
    population = fmode(population),
    land_area_skm = fmode(land_area_skm),
    timestamp = fmode(timestamp)
  ) %>%
  ungroup() %>%
  summarize(
    country = "World", iso3c = NA,
    population = sum(population, na.rm = TRUE),
    land_area_skm = sum(land_area_skm, na.rm = TRUE)
  ) %>%
  mutate(pop_density = population / land_area_skm)

add_world <- left_join(add_world1, add_world2, by = c("country", "iso3c"))
merged <- bind_rows(merged, add_world)

cv_data <- pivot_longer(merged,
  cols = c(
    "confirmed", "deaths", "recovered",
    "daily_confirmed", "daily_deaths", "daily_recovered"
  ),
  names_to = "variable", values_to = "cases") %>%
  arrange(country, variable, date) %>%
  rename(area = land_area_skm, density = pop_density) %>%
  mutate(rate = cases / population * 10^6) %>%
  ## new code to omit data before 2020-05-11
  filter(date < "2020-05-11")
## end of new code

cv_summary <- function(d, country_list = "World",
                       plot = TRUE, facet = "country",
                       status = c("confirmed", "deaths", "recovered")) {

  # based on `wes_palettes()` color schemes GrandBudapest1, IsleofDogs1,
  # IsleofDogs2 from the {wesanderson} package
  my_palette <- c(
    "#5B1A18", "#FD6467", "#F1BB7B", "#D67236",
    "#0F0D0E", "#9986A5", "#79402E", "#CCBA72", "#D9D0D3", "#8D8680",
    "#EAD3BF", "#AA9486", "#B6854D", "#39312F", "#1C1718"
  )

  if (facet == "country") {
    fill <- "variable"
    n <- length(unique(d$variable)) / 2
    # need only half of unique # of variable (3)
  }

  if (facet == "variable") {
    fill <- "country"
    n <- length(country_list)
    # need number of countries
  }

  if ("All" %in% country_list) {
    country_list <- unique(d$country)
    country_list <- setdiff(country_list, "World")
  }

  if ("World" %in% country_list) {
    d <- d %>% filter(country %in% country_list)

    totals <- d %>%
      group_by(variable) %>%
      summarize(
        country = "World",
        cases = max(cases),
        population = max(population),
        area = max(area),
        density = max(density),
        rate = max(rate, na.rm = TRUE),
        on = max(date)
      ) %>%
      select(country, variable, cases, population, area, density, rate, on) %>%
      arrange(variable) %>%
      ungroup()
  }

  if ("World" %nin% country_list) {
    d <- d %>% filter(country %in% country_list)
    totals <- d %>%
      group_by(country, variable) %>%
      summarize(
        cases = max(cases),
        population = max(population),
        area = max(area),
        density = max(density),
        rate = max(rate, na.rm = TRUE),
        on = max(date),
        gdp_capita = fmode(gdp_capita),
        income = fmode(income),
        life_expectancy = fmode(life_expectancy),
        max_sd = max(soc_dist),
        max_mr = max(mov_rest)
      ) %>%
      select(
        country, variable, cases, population, area, density, rate,
        gdp_capita, income, life_expectancy, max_sd, max_mr, on
      ) %>%
      arrange(country, variable) %>%
      ungroup()
  }

  if (plot == TRUE) {
    cc <- filter(d, variable %in% status)
    cum_cases_plot <- ggplot(
      data = cc,
      # use the tidy evaluation pronoun .data to slice the chosen fill
      # variable from the data frame
      aes(
        x = date, y = cases + 1, color = .data[[fill]],
        fill = .data[[fill]]
      )
    ) +
      geom_point(size = 0.5) +
      geom_line() +
      # use the tidy evaluation pronoun .data to slice the chosen facet_wrap
      # variable from the data frame
      facet_wrap(~ .data[[facet]], ncol = 5) +
      xlab("Date") +
      ylab("Log Cumulative Cases") +
      scale_y_log10(
        breaks = trans_breaks("log10", function(x) 10^x),
        labels = trans_format("log10", math_format(10^.x))
      ) +
      scale_color_manual(
        aesthetics = c("color", "fill"),
        name = NULL, values = my_palette
      )

    dc <- filter(d, variable %in% paste0("daily_", status))
    daily_cases_plot <- ggplot(
      data = dc,
      aes(
        x = date, y = cases, color = .data[[fill]],
        fill = .data[[fill]]
      )
    ) +
      geom_point(size = 0.5) +
      geom_line() +
      facet_wrap(~ .data[[facet]], ncol = 5) +
      xlab("Date") +
      ylab("Daily Cases") +
      scale_color_manual(
        aesthetics = c("color", "fill"),
        name = NULL, values = my_palette
      )
  }

  if (plot == TRUE) {
    return(list(
      totals = totals,
      cum_cases_plot = cum_cases_plot,
      daily_cases_plot = daily_cases_plot
    ))
  } else {
    return(list(totals = totals))
  }
}
```

# CHALLENGE 1
Plot global data on confirmed coronavirus infections, deaths, and recoveries. (Faceted by variable rather than country, although it doesn't really matter in this case.)
```{r}
cv_summary(cv_data, country_list = "World", plot = TRUE, facet = "variable")
```
  
# CHALLENGE 2
Plot data on confirmed coronavirus infections, deaths, and recoveries for G7 countries plus China, Russia, and Iran. Facet your plots first by “country” and then by “variable”.
```{r}
cv_summary(cv_data, country_list = c("United States", "United Kingdom", "Canada", "France", "Germany", "Italy", "Japan", "China", "Russia", "Iran"), plot = TRUE, facet = "country")
cv_summary(cv_data, country_list = c("United States", "United Kingdom", "Canada", "France", "Germany", "Italy", "Japan", "China", "Russia", "Iran"), plot = TRUE, facet = "variable")
```
  
# CHALLENGE 3
Summary data for ALL countries in the dataset, then filter dataset to only those countries with populations of over 1 million. How many countries does this tibble include?

## 930  
(But it was 918 then 924 as of May 14?)
```{r}
summ <- cv_summary(cv_data, country_list = "All", plot = FALSE)
d <- summ$totals # since tibble is outputted in list of one element
d <- filter(d, population > 1000000)
nrow(d)
```
  
# CHALLENGE 4
Which 10 countries have experienced the highest single-day rate of confirmed cases? See below:
```{r}
# OVERALL CONFIRMED CASES
overall <- filter(d, variable == "confirmed") %>% arrange(desc(rate))
## Top 10:
top_n(overall, 10, rate)$country
# DAILY CONFIRMED
daily <- filter(d, variable == "daily_confirmed", rate > 0) %>% arrange(desc(rate))
## Top 10:
top_n(daily, 10, rate)$country
```
  
# CHALLENGE 5
gdp_capita has a slope significantly different from 0 (P = 0.000326)  
Note: Density was significant when I filtered out 0's and used log() instead of log1p(), but it's p-value increased to 0.0866 here when I did it this way.
```{r}
# visualize distribution and scatterplots using {psych}
library(psych)
x <- select(overall, c("rate", "density","population","gdp_capita","income"))
pairs.panels(x[],
  smooth = FALSE,
  lm = TRUE,
  method = "pearson",
  hist.col = "#00AFBB",
  density = TRUE,
  ellipses = TRUE
)
# DAILY
y <- select(daily, c("rate", "density", "population", "gdp_capita", "income"))
pairs.panels(y[],
  smooth = FALSE,
  lm = TRUE,
  method = "pearson",
  hist.col = "#00AFBB",
  density = TRUE,
  ellipses = TRUE
)
# Continuous variables are not normally distributed, so log transform rate, density, population, gdp_capita in new tibble overall_log
overall_log <- tibble(rate = log1p(overall$rate), density = log1p(overall$density), population = log1p(overall$population), gdp_capita = log1p(overall$gdp_capita), income = overall$income)
# same for daily, which we'll use later
daily_log <- tibble(rate = log1p(daily$rate), density = log1p(daily$density), population = log1p(daily$population), gdp_capita = log1p(daily$gdp_capita), income = daily$income)
# visualize overall again - yay normal distribution!
pairs.panels(overall_log,
  smooth = FALSE,
  lm = TRUE,
  method = "pearson",
  hist.col = "#00AFBB",
  density = TRUE,
  ellipses = TRUE
)
# FULL MODEL USING OVERALL_LOG
m_full <- lm(rate ~ density + population + gdp_capita + income, data = overall_log)
summary(m_full)
detach(package:psych)
```
  
# CHALLENGE 6
## Best model: rate ~ density + gdp_capita  
(although AIC is not quite 2 less than density + population + gdp_capita)  
## Pseudo-R2 - Full: 0.55, Best: 0.54 
Which doesn't really make sense - shouldn't the best model have a higher value? Earlier on I got Full: 0.515, Best: 0.546, but then I guess I fiddled with filtering NA's and 0's and when I ran everything again got these results instead. Tried to backtrack and figure out what caused the difference, but haven't been able to replicate those orignal results.
```{r}
library(MASS)
library(MuMIn)
s <- stepAIC(m_full, scope = .~., direction = "both")
m_best <- lm(rate ~ density + gdp_capita, data = overall_log)
summary(m_best)
# Pseudo-R^2
r.squaredGLMM(m_full)
r.squaredGLMM(m_best)
```
Repeat this modeling process to evaluate what combination of explanatory variables best predict maximum daily infection rate. Are the important predictors the same? What additional or different variables are included?  
## Best model: rate ~ density + population + gdp_capita  
## Population size becomes an important predictor in daily infection rate.
```{r}
m_daily <- lm(rate ~ density + population + gdp_capita + income, data = daily_log)
(s <- stepAIC(m_daily, scope = .~., direction = "both"))
detach(package:MASS)
```
  
# CHALLENGE 7
Neither variable (max_sd or max_mr) improves the model significantly, although the delta_AICc of the model including max_mr and the model including max_sd are less than 2, making them viable alternatives.
```{r}
# check whether new variables should be transformed - they should
hist(daily$max_sd)
hist(daily$max_mr)
# generate possible models with addition of max_mr, max_sd, or both
daily_log <- add_column(daily_log, max_sd = log1p(daily$max_sd), max_mr = log1p(daily$max_mr))
best_daily <- lm(rate ~ density + population + gdp_capita, data = daily_log)
best_SD_MR <- lm(rate ~ density + population + gdp_capita + max_sd + max_mr, data = daily_log)
best_SD <- lm(rate ~ density + population + gdp_capita + max_sd, data = daily_log)
best_MR <- lm(rate ~ density + population + gdp_capita + max_mr, data = daily_log)
# AIC table to compare new models to original best
library(AICcmodavg)
(aic_table <- aictab(list(best_daily, best_SD_MR, best_SD, best_MR), modnames = c("best_daily", "best_SD_MR", "best_SD", "best_MR")))
# alternative(?): use add1() to see if p_values of max_sd or max_mr are significant, which they are not
add1(best_daily, scope = .~. + max_sd + max_mr, test = "F")
```
  
# CHALLENGE 8
The best mixed effects model is the full model, containing all 4 fixed effects (log(density), log(gdp_capita), soc_dist, mov_rest) along with random effects country and date.  
Its pseudo-R^2 is 0.519 (conditional - variance explained by fixed and random effects), and 0.279 (marginal - variance explained by fixed effects only).
```{r}
library(lme4)
# filter cv_data
cv_filtered <- filter(cv_data, population > 1000000, variable == "daily_confirmed", rate > 0)
# full model
m1 <- lmer(rate ~ log(density) + log(gdp_capita) + soc_dist + mov_rest + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
# 3 fixed effects
minusMR <- lmer(rate ~ log(density) + log(gdp_capita) + soc_dist + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
minusSD <- lmer(rate ~ log(density) + log(gdp_capita) + mov_rest + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
minusGDP <- lmer(rate ~ log(density) + soc_dist + mov_rest + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
minusDen <- lmer(rate ~ log(gdp_capita) + soc_dist + mov_rest + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
# 2 fixed effects
Den_GDP <- lmer(rate ~ log(density) + log(gdp_capita) + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
Den_SD <- lmer(rate ~ log(density) + soc_dist + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
Den_MR <- lmer(rate ~ log(density) + mov_rest + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
GDP_SD <- lmer(rate ~ log(gdp_capita) + soc_dist + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
GDP_MR <- lmer(rate ~ log(gdp_capita) + mov_rest + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
SD_MR <- lmer(rate ~ soc_dist + mov_rest + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
# 1 fixed effect
mDen <- lmer(rate ~ log(density) + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
mGDP <- lmer(rate ~ log(gdp_capita) + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
mSC <- lmer(rate ~ soc_dist + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
mMR <- lmer(rate ~ mov_rest + (1|country) + (1|date), data = cv_filtered, REML = FALSE)
# null model - random effects only
null <- lmer(rate ~ (1|country) + (1|date), data = cv_filtered, REML = FALSE)

# AIC table
library(AICcmodavg)
(aic_table <- aictab(list(m1, minusMR, minusSD, minusGDP, minusDen, Den_GDP, Den_SD, Den_MR, GDP_SD, GDP_MR, SD_MR, mDen, mGDP, mSC, mMR, null), modnames = c("full", "minusMR", "minusSD", "minusGDP", "minusDen", "Den_GDP", "Den_SD", "Den_MR", "GDP_SD", "GDP_MR", "SD_MR", "Den", "GDP", "SD", "MR", "null")
))

# Pseudo-R^2
r.squaredGLMM(m1)

detach(package:lme4)
detach(package:AICcmodavg)
detach(package:MuMIn)
```

